---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

First we unzip the file, and then load it to the activity variable:
```{r}
unzip("activity.zip")
activity <- read.csv("activity.csv",header=TRUE)
```

## What is mean total number of steps taken per day?

To calculate the total daily steps for each day, the aggregate function inserts the summarized data into the ActSum variable, and then we plot the daily steps histogram:
```{r}
ActSum <- aggregate(activity$steps,by=list(activity$date),FUN=sum)
hist(ActSum$x,col="lightblue",xlab = "Number of Steps",main="Daily Steps Histogram")
```

```{r}
dailyMean <- format(mean(ActSum$x,na.rm=TRUE),digits=2,nsmall = 2,big.mark = ",", big.interval=3)
dailyMedian <- format(median(ActSum$x,na.rm=TRUE),digits=2,nsmall = 2,big.mark = ",", big.interval = 3)
```

The daily mean is `r dailyMean`, the daily median is `r dailyMedian`

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
